Abstractions for Type-Level Programming

### 2022-03-22 slides.tex
Slides

\usepackage{ulem}
\usepackage{xspace}
\usepackage{mathtools}
\def\A{\normalfont\text{A}}
\def\B{\normalfont\text{B}}
\def\C{\normalfont\text{C}}
\def\D{\normalfont\text{D}}
\def\T{\normalfont\text{T}}
\def\<:{\allowbreak\operatorname{\normalfont\text{<}:}\allowbreak}
\def\Fsub{\normalfont{F}\ensuremath{_{\<:}}\xspace}

@@

The regex library uses technology developed in this thesis.
\end{slide}

@@

\begin{slide}{Timeline of my PhD}
\timeline{0}{0}{0}{0}{0}
\end{slide}

\begin{slide}{Part I: Generalized Singleton Types}
\timeline{80}{0}{80}{80}{80}
\end{slide}

@@

\begin{slide}{Generalized, or not singleton at all?}
\begin{code}
dependent
def int2str(x: Int)
  : { if (x @\tp{== 0}@) "zero" @\tp{else}@ (_: String) }
  =   if (x == 0) "zero" else readString()
\end{code}
\begin{overprint}
\onslide<1>
\begin{block}{Definition}
A singleton type is a type that contains exactly one value.
\end{block}

\onslide<2>
\begin{block}{Definition}
\sout{A singleton type is a type that contains exactly one value.}
\end{block}
\end{overprint}
\end{slide}

@@

\begin{slide}{Part II: Match Types}
\timeline{80}{80}{0}{80}{80}
\end{slide}

@@

\begin{overprint}
\onslide<2>
Match types can be explained in terms of generalized singletons:

@@

\onslide<4>
The type/term correspondence dictates match type semantics:

@@

@@

elem[Any]("foo")
\end{code}
\vspace{-10pt}
\end{overprint}
\end{slide}

@@

@@

We cannot prove disjointness when variance is involved:

>>>>

\usepackage[sorting=none, backend=biber, maxbibnames=9]{biblatex}
\settowidth\labelnumberwidth{\usebeamertemplate*{bibliography item}}
\AtEveryBibitem{\clearfield{note}}
\AtEveryBibitem{\clearfield{volume}}
\AtBeginBibliography{\small}
\bibliography{bibliography}

\usepackage{textpos}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{mathtools}
\def\A{\normalfont\text{A}}
\def\B{\normalfont\text{B}}
\def\C{\normalfont\text{C}}
\def\D{\normalfont\text{D}}
\def\T{\normalfont\text{T}}
\def\<:{\allowbreak\operatorname{\normalfont\text{<}:}\allowbreak}
\def\Fsub{\normalfont{F}\ensuremath{_{\<:}}\xspace}

@@

The regex library uses a first-class language construct:
\begin{itemize}
  \item easier to use
  \item faster to compile
\end{itemize}

@@

\begin{slide}{Timeline of my PhD}
\begin{overprint}
\onslide<1>
\timeline{0}{0}{0}{0}{0}
\onslide<2>
\timeline{80}{80}{80}{0}{80}
\vspace{-10pt}
\begin{refsection}%
\raggedright
\phantom{\cite{blanvillain2022regular}}%
\printbibliography[heading=subbibliography]%
\end{refsection}
\vspace{-40pt}
\onslide<3>
\timeline{80}{0}{80}{80}{80}
\onslide<4>
\timeline{80}{0}{80}{80}{80}
\vspace{-10pt}
\begin{refsection}%
\phantom{\cite{schmid2020coming}}%
\printbibliography[heading=subbibliography]%
\end{refsection}
\vspace{-40pt}
\end{overprint}
\end{slide}

@@

@@

\begin{slide}{Part II: Match Types}
\begin{overprint}
\onslide<1>
\timeline{80}{80}{0}{80}{80}
\onslide<2>
\timeline{80}{80}{0}{80}{80}
\vspace{-10pt}
\begin{refsection}%
\phantom{\cite{blanvillain2022type}}%
\printbibliography[heading=subbibliography]%
\end{refsection}
\vspace{-40pt}
\end{overprint}
\end{slide}

@@

\begin{overprint}
\onslide<2>
Some examples of reduction:
\begin{itemize}
  \item |Elem[String] =:= Char|
  \item |Elem[Int] =:= Int|
  \item |Elem[List[Int]] =:= Int|
  \item |Elem[Any] =:= Elem[Any]|
\end{itemize}

\onslide<4>
Match types can be explained in terms of generalized singletons:

@@

@@

\onslide<6>
The type/term correspondence dictates match type semantics:

@@

elem[Any]("foo")
\end{code}
\end{overprint}
\end{slide}

@@

\begin{slide}{Formalization vs implementation}
Our implementation goes much further than System FM:
\begin{itemize}
  \item recursion / non-termination
  \item type parameter bindings (|case List[t]|)
  \item empty types
  \item variance
\end{itemize}
\end{slide}

@@

We cannot prove disjointness when variance is involved, |F[T1]| and |F[T2]| always overlap:
